### 2013-03-11 doc/Version_Control/Git-SVN_Quick_Reference.tex
Explanations on how to use Git in addition to SVN. Dirty and elementary. 	new file:   doc/Version_Control/Git-SVN_Quick_Reference.tex

\documentclass[11pt,twoside]{article}

\begin{document}

\title{GIT-SVN basic commands}
\maketitle

\subsection*{Important remark}

It might be useful to issue some git-svn commands beforehand with the \verb|-n| option (or with its longer name \verb|--dry-run|).
This option shows what will be impacted by the \texttt{git-svn} command before issuing the actual command.
This can be used with the \texttt{dcommit}, \texttt{rebase}, \texttt{branch} and \texttt{tag} commands.

\subsection*{Git-svn -- Initial checkout}
\begin{verbatim}
$ git svn clone -s \
  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D_GLOBE
\end{verbatim}
The previous command checkouts the full svn history and converts it into a git history. It takes a huge amount of time. If the full history is not needed, one can checkout a partial history.
For instance, checking out revisions from $20\:000$ to the last one (\texttt{HEAD}) is done with:
\begin{verbatim}
$ git svn clone -s -r 20000:HEAD \
  svn+ssh://svn@geodynamics.org/cig/seismo/3D/SPECFEM3D\_GLOBE
\end{verbatim}

\subsection*{Git-svn -- SVN branches}

Taking a look at the different branches available:
\begin{verbatim}
$ git branch -a
* master
  remotes/NOISE_TOMOGRAPHY
  remotes/ORIGINAL
  remotes/SPECFEM3D_GLOBE_SUNFLOWER
  remotes/adjoint
  remotes/pluggable
  remotes/tags/v5.1.5
  remotes/trunk
\end{verbatim}

Creating a new branch \texttt{SPECFEM3D\_GLOBE\_ADIOS} on the svn server:
\begin{verbatim}
$ git svn branch SPECFEM3D_GLOBE_ADIOS
\end{verbatim}

Switching to a remote branch. It is mandatory, else the commit will be done in the master branch
\begin{verbatim}
$ git checkout -b local/SPECFEM3D_GLOBE_ADIOS \
  SPECFEM3D_GLOBE_ADIOS
\end{verbatim}

\subsection*{Git -- Checking the status of the file.}
Before doing any modification it is mandatory to know which files are tracked, have been modified or deleted.

\begin{verbatim}
$ git status
\end{verbatim}

If some files appear to be untracked they can be added to the local git repository with:

\begin{verbatim}
$ git add filename
\end{verbatim}

\begin{verbatim}
$ git status 
# should now return the added file
# as 'new file' unstead of 'untracked'.
\end{verbatim}

If some files were modified, before committing the modifications, it may be useful to see what has changed :

\begin{verbatim}
$ git diff filename
\end{verbatim}

\subsection*{Git -- Committing changes into the local repository}

\begin{verbatim}
$ git commit
\end{verbatim}

This command will open your favorite text editor (it is likely to be vi without any particular configuration) and ask you for a comment about the commit. If nothing is written, the commit will abort.

\subsection*{Git -- Removing files}

\begin{verbatim}
$ rm filename # remove the file from your workspace.
\end{verbatim}
\begin{verbatim}
$ git rm filename 
# prepare to remove the file from the git local repository.
\end{verbatim}
\begin{verbatim}
$ git commit #... 
\end{verbatim}

\subsection*{Git -- Moving files}

\begin{verbatim}
$ git mv old_file new_file
\end{verbatim}
\begin{verbatim}
$ git commit
\end{verbatim}

OR

\begin{verbatim}
$ mv old_file new_file
\end{verbatim}
\begin{verbatim}
$ git rm old_file
\end{verbatim}
\begin{verbatim}
$ git add new_file
\end{verbatim}
\begin{verbatim}
$ git commit
\end{verbatim}

\subsection*{Git -- Commit history}

\begin{verbatim}
$ git log
\end{verbatim}

\subsection*{Git -- Changing the last commit}

\begin{verbatim}
$ git commit -m 'initial commit'
\end{verbatim}
\begin{verbatim}
$ git add forgotten_file
\end{verbatim}
\begin{verbatim}
$ git commit --amend
\end{verbatim}
\subsection*{Git -- Unstaging a staged file (i.e. Removing a file from the list of files to be 
committed)}

\begin{verbatim}
$ git status 
# the file has to be in the 'changes to be committed' section
\end{verbatim}

\begin{verbatim}
$ git reset HEAD filename
filename: locally modified
\end{verbatim}
\begin{verbatim}
$ git status  
# the file should be in the 'Changes not staged for commit' section
\end{verbatim}

\subsection*{Git -- Unmodifying a Modified File}

\begin{verbatim}
$ git checkout -- filename
\end{verbatim}
\begin{verbatim}
$ git status
\end{verbatim}

\subsection*{Git-svn -- Back to the svn server}

Push modifications up to the svn server:
\begin{verbatim}
$ git svn dcommit
\end{verbatim}
That will also bring your local tree up to date. To bring your tree up to date in general, run: 
\begin{verbatim}
$ git svn rebase
\end{verbatim}
This will update your local checkout and then re-apply your local un-submitted commits on top of the new trunk. 
If you want to get the commits for all branches that exist in your clone: 
\begin{verbatim}
$ git svn fetch
\end{verbatim}

\subsection*{References and usefull readings}

\verb|http://git-scm.com/book|
\\
\verb|http://git.or.cz/course/svn.html|
\\
\verb|http://trac.parrot.org/parrot/wiki/git-svn-tutorial|

\end{document}