### 2014-11-12 utils/ADJOINT_TOMOGRAPHY_TOOLS/iterate_adj/model_plot/gmt/xc_and_seis.tex
move ADJOINT_TOMOGRAPHY_TOOLS/ to utils/ to reflect the developmental (or obsolete) state of these tools; other efforts are ongoing outside of SPECFEM3D; see also src/tomography/

\documentclass[pdf,mpa]{prosper}
\usepackage{portland}
\begin{document}
\begin{slide}{}
\vspace{-1.7cm}
\fcolorbox{white}{white}{
\hspace{-1cm} 
\begin{tabular}{c}
\includegraphics[width=4.5cm,angle=-90]{/home/carltape/ADJOINT_TOMO/ADJOINT_TOMO_OUTPUT/model_plot_gmt_OUTPUT/VERT_01/vert_01_xc_LGU_CI_206_12659440_vs_m16_m00_077.eps}
\\
\includegraphics[width=3.8cm,angle=-90]{/home/carltape/ADJOINT_TOMO/ADJOINT_TOMO_OUTPUT/misfit_plot_OUTPUT/T006_T030_win1_1D1/12659440_T006_T030_LGU_CI_m16_seis_3.eps}
\end{tabular}
}
\end{slide}
\end{document}