### 2015-04-24 doc/USER_MANUAL/14_post_processing.tex

\chapter{Post-Processing Scripts}

Several post-processing scripts/programs are provided in the \texttt{utils/}
directory, and most of them need to be adjusted when used on different
systems, for example, the path of the executable programs. Here we
only list a few of the available scripts and provide a brief description,
and you can either refer to the related sections for detailed usage
or, in a lot of cases, type the script/program name without arguments
for its usage.

\section{Process Data and Synthetics}\label{sec:Process-data-and-syn}

In many cases, the SEM synthetics are calculated and compared to data
seismograms recorded at seismic stations. Since the SEM synthetics
are accurate for a certain frequency range, both the original data
and the synthetics need to be processed before a comparison can be
made.

For such comparisons, the following steps are recommended:
\begin{enumerate}
\item Make sure that both synthetic and observed seismograms have the correct
station/event and timing information.
\item Convolve synthetic seismograms with a source time function with the
half duration specified in the \texttt{CMTSOLUTION} file, provided,
as recommended, you used a zero half duration in the SEM simulations.
\item Resample both observed and synthetic seismograms to a common sampling
rate.
\item Cut the records using the same window.
\item Remove the trend and mean from the records and taper them.
\item Remove the instrument response from the observed seismograms (recommended)
or convolve the synthetic seismograms with the instrument response.
\item Make sure that you apply the same filters to both observed and synthetic
seismograms. Preferably, avoid filtering your records more than once.
\item Now, you are ready to compare your synthetic and observed seismograms.
\end{enumerate}
\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
directory:

\subsection{Data processing script \texttt{process\_data.pl}}

This script cuts a given portion of the original data, filters it,
transfers the data into a displacement record, and picks the first
P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
without any argument. An example of the usage of the script:
\begin{lyxcode}
{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
\end{lyxcode}
\noindent which has resampled the SAC files to a sampling rate of
1 seconds, cut them between 0 and 4000 seconds, transfered them into
displacement records and filtered them between 40 and 500 seconds,
picked the first P and S arrivals, and added suffix `\texttt{bp}'
to the file names.

Note that all of the scripts in this section actually use the SAC
and/or IASP91 to do the core operations; therefore make sure that
the SAC and IASP91 packages are installed properly on your system,
and that all the environment variables are set properly before running
these scripts.

\subsection{Synthetics processing script \texttt{process\_syn.pl}}

This script converts the synthetic output from the SEM code from ASCII
to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
An example of the usage of the script:
\begin{lyxcode}
{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
\end{lyxcode}
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample
the SAC files with a sampling rate of 1 seconds, cut them between
0 and 4000 seconds, filter them between 40 and 500 seconds with the
same filter used for the observed data, pick the first P and S arrivals,
and add the suffix `\texttt{bp}' to the file names.

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source time function with a given half duration, etc.
Type \texttt{process\_syn.pl} without any argument for a detailed
usage.

In order to convert between SAC format and ASCII files, useful scripts
are provided in the subdirectories ~\\
 \texttt{utils/sac2000\_alpha\_convert/} and \texttt{utils/seis\_process/asc2sac/}.

\subsection{Script \texttt{rotate.pl}}

The original data and synthetics have three components: vertical (BHZ
resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
for most seismology applications, transverse and radial components
are also desirable. Therefore, we need to rotate the horizontal components
of both the data and the synthetics to the transverse and radial direction,
and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
\begin{lyxcode}
rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~

rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
\end{lyxcode}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

\section{Collect Synthetic Seismograms}

The forward and adjoint simulations generate synthetic seismograms
in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
simulation, the files are named like \texttt{NT.STA.BX?.semd} for
two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
SAC format, where NT and STA are the network code and station name,
and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
and \ref{cha:channel-codes} for further details.

The adjont simulations generate synthetic seismograms with the name
\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
for details). The kernel simulations output the back-reconstructed
synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
for the purpose of checking the accuracy of the reconstruction. Refer
to Section \ref{sec:Adjoint-simulation-finite} for further details.

You do have further options to change this default output behavior,
given in the main constants file \texttt{constants.h} located in \texttt{src/shared/}
directory:
\begin{description}
\item [{\texttt{SEISMOGRAMS\_BINARY}}] set to \texttt{.true.} to have seismograms
written out in binary format.
\item [{\texttt{WRITE\_SEISMOGRAMS\_BY\_MASTER}}] Set to \texttt{.true.}
to have only the master process writing out seismograms. This can
be useful on a cluster, where only the master process node has access
to the output directory.
\item [{\texttt{USE\_OUTPUT\_FILES\_PATH}}] Set to \texttt{.false.} to
have the seismograms output to \texttt{LOCAL\_PATH} directory specified
in the main parameter file \texttt{DATA/Par\_file}. In this case,
you could collect the synthetics onto the frontend using the \texttt{collect\_seismo\_lsf\_multi.pl}
script located in the \texttt{utils/Cluster/lsf/} directory. The usage
of the script would be e.g.:

\begin{lyxcode}
collect\_seismo.pl~machines~DATA/Par\_file
\end{lyxcode}
\end{description}
where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
the parameter file used to extract the \texttt{LOCAL\_PATH} directory
used for the simulation.

\section{Clean Local Database}

After all the simulations are done, the seismograms are collected,
and the useful database files are copied to the frontend, you may
need to clean the local scratch disk for the next simulation. This
is especially important in the case of kernel simulation, where very
large files are generated for the absorbing boundaries to help with
the reconstruction of the regular forward wavefield. A sample script
is provided in \texttt{utils/}:
\begin{lyxcode}
cleanbase.pl~machines
\end{lyxcode}
where \texttt{machines} is a file containing the node names.

\section{Plot Movie Snapshots and Synthetic Shakemaps}

\subsection{Script \texttt{movie2gif.gmt.pl}}

With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
code to convert these binary movie data into GMT xyz files for futher
processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
is provided to do this conversion, and then plot the movie snapshots
in GMT, for example:}{\small \par}
\begin{lyxcode}
movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
\end{lyxcode}
which for the first through the 40th movie frame, converts the \texttt{moviedata}
files into GMT xyz files, interpolates them using the 'nearneighbor'
command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
and `\texttt{-p}' are both optional.

\subsection{Script \texttt{plot\_shakemap.gmt.pl}}

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation ~\\
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
code to convert the binary shakemap data into GMT xyz files. A sample
script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
and then plot the shakemaps in GMT, for example:
\begin{lyxcode}
plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
\end{lyxcode}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

\section{Map Local Database}

A sample program \texttt{remap\_database} is provided to map the local
database from a set of machines to another set of machines. This is
especially useful when you want to run mesher and solver, or different
types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
\begin{lyxcode}
run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
\end{lyxcode}
where \texttt{old\_machines} is the LSF machine file used in the previous
simulation, and \texttt{150} is the number of processors in total.