### 2015-06-22 doc/USER_MANUAL/10_gravity_calculations.tex

\chapter{Gravity integral calculations for the gravity field of an Earth model}

SPECFEM can now compute the gravity field as well as its derivatives (i.e., gravity gradiometry)
generated by any given 3D Earth model at a given elevation, i.e., on a given observation surface.
That feature is still experimental but should work just fine.\\

To see how it is implemented and to use it, type this in the root directory of the code:
\begin{verbatim}
grep -i GRAVITY_INTEGRALS src/*/*
\end{verbatim}