### 2015-06-22 doc/USER_MANUAL/sponsors.tex

\section*{Current and past main participants or main sponsors of the SPECFEM project (in no particular order)}
%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.125\textwidth]{figures/logo_cnrs.png}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_princeton.jpg}\vspace*{2truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_aix_marseille_universite.png}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_ETH.jpg}\vspace*{2truemm}
\includegraphics[width=0.12\textwidth]{figures/logo_CSC_China.jpg}\vspace*{0.02truemm}
\includegraphics[width=0.15\textwidth]{figures/logo_inria.jpg}\vspace*{2truemm}
\includegraphics[width=0.14\textwidth]{figures/logo_UPPA.png}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.15\textwidth]{figures/logo_NVIDIA.jpg}\vspace*{2truemm}
\includegraphics[width=0.125\textwidth]{figures/logo_IUF.jpg}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Caltech.png}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_Harvard.jpg}\vspace*{2truemm}
\includegraphics[width=0.135\textwidth]{figures/logo_IPGP.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_ANR.png}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_NSF.png}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.190\textwidth]{figures/logo_European_Union.png}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_GENCI.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_PRACE.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_CINES.png}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Oak_Ridge.png}\vspace*{2truemm}
\hspace*{3mm}\includegraphics[width=0.130\textwidth]{figures/logo_fondation_Del_Duca.png}
\includegraphics[width=0.130\textwidth]{figures/logo_CIG.png}\vspace*{2truemm}
\par\end{centering}
%
\vspace*{-2truemm}
%
\noindent \begin{centering}
\includegraphics[width=0.112\textwidth]{figures/logo_University_of_Toronto.jpg}\vspace*{2truemm}
\includegraphics[width=0.109\textwidth]{figures/logo_INGV.jpg}\vspace*{2truemm}
\includegraphics[width=0.102\textwidth]{figures/logo_Univ_Toulouse.jpg}\vspace*{2truemm}
\includegraphics[width=0.112\textwidth]{figures/logo_TOTAL.jpg}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Fairbanks.jpg}\vspace*{2truemm}
\includegraphics[width=0.120\textwidth]{figures/logo_CINECA.jpg}\vspace*{2truemm}
\includegraphics[width=0.140\textwidth]{figures/logo_Intel_Exascale_Labs.png}\vspace*{2truemm}
\includegraphics[width=0.130\textwidth]{figures/logo_Maison_Simulation.jpg}
\par\end{centering}
\end{figure}