### 2015-10-12 doc/USER_MANUAL/figures/tomo_dir_struct.tex

\documentclass{standalone}

\usepackage{amsmath}
\usepackage{pgfplots}
\pgfplotsset{compat=1.3}

% Latex generated graphics
\usepackage{tikz}

% tikz libraries
\usetikzlibrary{trees}

%\usetikzlibrary{positioning}
%\usetikzlibrary{shapes}
%\usetikzlibrary{calc}
%\usetikzlibrary{backgrounds}
%\usetikzlibrary{fit}
%\usetikzlibrary{patterns}

%\tikzset{timed/.style=dashed}

% colors
\usepackage{color}
\definecolor{myred}{rgb}{0.5,0.2,0.2}

\begin{document}

\tikzstyle{every node}=[draw=black,thick,anchor=west,inner sep=2pt,minimum size=1pt]

\tikzstyle{links}=[draw=red,fill=myred!20]
%\tikzstyle{file}=[dashed,fill=gray!50]
\tikzstyle{file}=[draw=none]
\tikzstyle{missing}=[draw=white]

\begin{tikzpicture}[%
  grow via three points={one child at (0.5,-0.7) and
  two children at (0.5,-0.7) and (0.5,-1.4)},
  edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)}
  ]
  \node {TOMO\_ROOT\_DIR}
    child { node {bin}}
    child { node [file] {kernels\_list.txt}}
    child { node {INPUT\_KERNELS}
        child { node [links] {event1\_DATABASES\_MPI}
          child { node [file] {proc***\_alpha\_kernel.bin}}
          child { node [file] {proc***\_beta\_kernel.bin}}
          child { node [file] {proc***\_rho\_kernel.bin}}
          child { node [file] {proc***\_hess\_kernel.bin}}
      }
     child [missing] {}
     child [missing] {}
     child [missing] {}
     child [missing] {}
     child { node [links] {event2\_DATABASES\_MPI}
       child { node [file] {....}}
      }
      child [missing] {}
      child { node [links] {....}}
    }
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child [missing] {}
    child { node {OUTPUT\_SUM}};
\end{tikzpicture}

\end{document}