### 2017-06-22 doc/USER_MANUAL/08b_performing_full_waveform_inversion_FWI_or_source_inversions.tex
fixed a typo in the users manual

\chapter{Performing full waveform inversion (FWI) or source inversions}\label{cha:FWI}

\red{\Large ANNOUNCEMENT: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain. For FWI inversions for the model, also refer to the new examples provided in
directory \texttt{EXAMPLES}.}