### 2017-11-13 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

% !TEX encoding = IsoLatin
%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english]{book}
\usepackage[T1]{fontenc}
%% DK DK we do not accept accented letters, use the old LaTeX syntax
%% DK DK for them instead: \'e \`e  \"e \^e and so on.
%% DK DK \usepackage[latin1]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

% table of contents and numbering
\usepackage{tocbibind}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\usepackage{babel}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{amssymb}

\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}

% biblio
\usepackage[authoryear]{natbib}
% biblio GJI
\bibliographystyle{abbrvnat}

% fonts
\usepackage{times}

% figures
\usepackage{graphicx}

% we are running pdflatex, so convert .eps files to .pdf
%\usepackage[dvips]{epsfig}

% colors to show the corrections
\usepackage[dvipsnames,usenames]{xcolor}
\definecolor{light-gray}{rgb}{0.7,0.7,0.7}
\definecolor{dark-gray}{rgb}{0.1,0.1,0.3}
\definecolor{light-blue}{rgb}{0.7,0.7,0.9}
\definecolor{dark-blue}{rgb}{0.2,0.2,0.5}

% hyperlinks to sections and references
\usepackage[pdftex]{hyperref}
% more options
\hypersetup{
  bookmarks=true,
  bookmarksnumbered=true,
  pdfpagemode=UseNone,
  pdfstartview=FitH,
  pdfpagelayout=SinglePage,
  pdfborder={0 0 0},
  colorlinks=true,      % color links
  citecolor=dark-gray,  % citation links
  filecolor=dark-gray,  % file links
  linkcolor=dark-gray,  % internal link (rgb values)
  urlcolor=dark-blue    % external weblinks
}

% create thumbnails
%\usepackage[pdftex]{thumbpdf}

\usepackage{breakurl}

% BibTeX logo
\newcommand{\BibTeX}{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em{T}\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\newcommand{\noun}[1]{\textsc{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxcode}
{\par\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\raggedright
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}%
 \item[]}
{\end{list}}
\newenvironment{lyxlist}[1]
{\begin{list}{}
{\settowidth{\labelwidth}{#1}
 \setlength{\leftmargin}{\labelwidth}
 \addtolength{\leftmargin}{\labelsep}
 \renewcommand{\makelabel}[1]{##1\hfil}}}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%\renewcommand{\baselinestretch}{1.5}

% hyperlinks to sections and references
\newcommand{\urlwithparentheses}[1]{(\url{#1})}

\newcommand{\nexxi}{\mbox{\texttt{NEX\_XI\/}}}
\newcommand{\nexeta}{\mbox{\texttt{NEX\_ETA\/}}}
\newcommand{\nprocxi}{\mbox{\texttt{NPROC\_XI\/}}}
\newcommand{\nproceta}{\mbox{\texttt{NPROC\_ETA\/}}}
\newcommand{\nchunks}{\mbox{\texttt{NCHUNKS\/}}}
\newcommand{\eg}{e.g.,}

% colors for the text
\newcommand{\red}[1]{\textbf{\textcolor{Red}{#1}}}

% some of Dimitri's aliases for math symbols
\newcommand{\Grad}{\mbox{\boldmath $\bf \nabla$}}
\newcommand{\Div}{\Grad \cdot }
\newcommand{\Laplacian}{\nabla^2}
\newcommand{\pressure}{p}
\newcommand{\Int}{\displaystyle\int}
\newcommand{\Sum}{\displaystyle\sum}
\renewcommand{\epsilon}{\varepsilon}
\newcommand{\myiomega}{\mathbf{i} \, \omega}
\newcommand{\p}{\partial}
\newcommand{\boldu}{\mathbf{u}}
\newcommand{\boldv}{\mathbf{v}}
\newcommand{\boldw}{\mathbf{w}}
\newcommand{\boldx}{\mathbf{x}}
\newcommand{\boldy}{\mathbf{y}}
\newcommand{\boldm}{\mathbf{m}}
\newcommand{\boldp}{\mathbf{p}}
\newcommand{\boldg}{\mathbf{g}}
\newcommand{\bolds}{\mathbf{s}}
\newcommand{\boldn}{\mathbf{\hat{n}}}
\newcommand{\boldd}{\mathbf{d}}
\newcommand{\boldf}{\mathbf{f}}
\newcommand{\boldB}{\mathbf{B}}
\newcommand{\boldc}{\mathbf{c}}
\newcommand{\boldC}{\mathbf{C}}
\newcommand{\boldE}{\mathbf{E}}
\newcommand{\boldH}{\mathbf{H}}
\newcommand{\boldI}{\mathbf{I}}
\newcommand{\boldsigma}{\mbox{\boldmath $\bf \sigma$}}
\newcommand{\boldxi}{\mbox{\boldmath $\bf \xi$}}
\newcommand{\boldvarepsilon}{\mbox{\boldmath $\bf \varepsilon$}}
\newcommand{\bolddelta}{\mbox{\boldmath $\bf \delta$}}
\newcommand{\boldtau}{\mbox{\boldmath $\bf \tau$}}
\newcommand{\Boxmath}{\Lambda}
\newcommand{\rmd}{\, \mathrm{d}}
\newcommand{\pv}{\ ,}
\newcommand{\pp}{\ .}

\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Start of document.

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\thispagestyle{empty}\vspace*{-1.8truecm} %
\makebox[1\textwidth]{%
\includegraphics[width=0.83\paperwidth]{figures/specfem_3d_Cartesian-cover.pdf} %
}
\par\end{center}

\title{\textbf{SPECFEM3D}\\
\textbf{Cartesian}\\
 \textbf{User Manual}}

\author{\copyright \, CNRS (France), Princeton University (USA),\\
and ETH Z\"urich (Switzerland)\\
Version 3.0}

% date of last edit
\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\include{authors}

\newpage{}

\include{sponsors}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Table of contents
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage{}

\tableofcontents{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\include{01_introduction}

\include{02_getting_started}

\include{03_mesh_generation}

\include{04_creating_databases}

\include{05_running_the_solver}

\include{06_fault_sources}

\include{07_adjoint_simulations}

\include{08_doing_tomography}

\include{08b_performing_full_waveform_inversion_FWI_or_source_inversions}

\include{09_noise_simulations}

\include{10_gravity_calculations}

\include{11_graphics}

\include{12_running_scheduler}

\include{13_changing_the_model}

\include{14_post_processing}

\include{15_informations_for_developers}

\include{bug_reports}

\include{notes_and_acknowledgement}

\include{copyright_and_version}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
%dummy comment inserted by tex2lyx to ensure that this paragraph is not empty

\include{A_reference_frame}

\include{B_channel_codes}

\include{C_troubleshooting}

\include{D_license}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}