### 2023-04-23 doc/USER_MANUAL/15_informations_for_developers.tex

\chapter{Information for developers of the code, and for people who want to learn how the technique works}\label{cha:developers}

You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):
\begin{verbatim}
git clone --recursive https://github.com/SPECFEM/specfem1d.git
\end{verbatim}

\noindent
We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\newline
\texttt{utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn} of the specfem3d package.
They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
Also useful to test new ideas by modifying these simple codes to run some tests.
We also have a similar, even simpler, demo source code for the 2D case in directory\newline
\texttt{utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn} of the specfem2d package.\newline

\noindent
For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
official package, see \url{https://github.com/SPECFEM/specfem3d/wiki} .