### 2023-06-02 doc/USER_MANUAL/notes_and_acknowledgement.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{Notes \& Acknowledgments}
\addcontentsline{toc}{chapter}{Notes \& Acknowledgments}

In order to keep the software package thread-safe in case a multithreaded
implementation of MPI is used, developers should not add modules or
common blocks to the source code but rather use regular subroutine
arguments (which can be grouped in ``derived types'' if needed for
clarity).\newline

The Gauss-Lobatto-Legendre subroutines in \texttt{gll\_library.f90}
are based in part on software libraries from the Massachusetts Institute
of Technology, Department of Mechanical Engineering (Cambridge, Massachusetts,
USA). The non-structured global numbering software was provided by
Paul F. Fischer (Brown University, Providence, Rhode Island, USA,
now at Argonne National Laboratory, USA).\newline

\href{http://www.opendx.org}{OpenDX} is open-source
based on IBM Data Explorer, \href{http://www.avs.com}{AVS}
is a trademark of Advanced Visualization Systems, and \href{http://www.paraview.org}{ParaView}
is an open-source visualization platform.\newline

Please provide your feedback, questions, comments, and suggestions through the github repository.